---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ID and response

* 'id': unique identifier for the customer
* 'attrition_flag': whether the customer is churned (0 = no; 1 = yes)

##  Categorical

* 'gender': gender of the customer
* 'education_level': education level of the customer
* 'income_category': income range of the customer

# Numeric

* 'total_relationship_count': number of relationships
* 'months_inactive_12_mon': number of months the customer is inactive in the past 12 months
* 'credit_limit': the customer's credit limit
* 'total_revolving_bal': the customer's total revolving balance
* 'total_amt_chng_q4_q1': the amount the balance changed from Q4 to Q1
* 'total_trans_amt': the value of all the customer's transactions in the period
* 'total_trans_ct': the number of all of the customer's transactions
* 'total_ct_chng_q4_q1': the difference in number of the customer's transactions from Q4 to Q1
* 'avg_utilization_ratio': the customer's average utilization ratio during the period

Strategy:

* change attrition flag to "churned" yes/no
* create summarize_churn function early on
* for linear, log on credit limit (maybe others?)
* Lot of numeric, so xgboost early on
* Try turning income category and education level into ordinal numbers (unknown as either 0 or as mean impute)

EDA:

* Density of all numeric, comparing attrition and not
* AUC of attrition vs all predictors

* Relationship between relationships/months inactive (line plot)
* Credit limit, age, total revolving balance, total amt change, (density plot or bucketed line plot)
* gender, Education + income and churn (bar plot)
* Total relationships and churn (line plot)

```{r}
library(tidymodels)
library(tidyverse)
library(scales)
library(lubridate)
library(stacks)
theme_set(theme_light())

doParallel::registerDoParallel(cores = 4)
```

```{r}
education_levels <- c("Unknown", "Uneducated", "High School", "College", "Graduate", "Doctorate", "Post-Graduate")
```

```{r}
process_data <- function(tbl) {
  tbl %>%
    mutate(education_level = fct_relevel(education_level, education_levels),
           income_category = fct_reorder(income_category, parse_number(income_category)),
           income_category = fct_relevel(income_category, "Less than $40K"))
}
```


```{r}
dataset <- read_csv("~/Downloads/sliced-s01e07-HmPsw2/train.csv") %>%
  mutate(churned = ifelse(attrition_flag == 1, "yes", "no")) %>%
  select(-attrition_flag) %>%
  process_data()
holdout <- read_csv("~/Downloads/sliced-s01e07-HmPsw2/test.csv") %>%
  process_data()

set.seed(2021)
spl <- initial_split(dataset)
train <- training(spl)
test <- testing(spl)

train_5fold <- train %>%
  vfold_cv(5)
```

```{r}
summarize_churn <- function(tbl) {
  tbl %>%
    summarize(n = n(),
              n_churned = sum(churned == "yes"),
              pct_churned = n_churned / n,
              low = qbeta(.025, n_churned + .5, n - n_churned + .5),
              high = qbeta(.975, n_churned + .5, n - n_churned + .5)) %>%
    arrange(desc(n))
}

plot_categorical <- function(tbl, category, ...) {
  if (!is.factor(pull(tbl, {{ category }}))) {
    tbl <- tbl %>%
      mutate({{ category }} := fct_reorder({{ category }}, pct_churned))
  }
  
  tbl %>%
    ggplot(aes(pct_churned, {{ category }}, ...)) +
    geom_col(position = position_dodge()) +
    geom_errorbarh(position = position_dodge(width = 1), aes(xmin = low, xmax = high), height = .2) +
    scale_x_continuous(labels = percent) +
    labs(x = "% in category that churned")
}

train %>%
  group_by(gender) %>%
  summarize_churn() %>%
  plot_categorical(gender)

train %>%
  group_by(education_level, gender) %>%
  summarize_churn() %>%
  plot_categorical(education_level, fill = gender, group = gender)

train %>%
  group_by(education_level) %>%
  summarize_churn() %>%
  plot_categorical(education_level) +
  coord_flip() +
  labs(x = "% churned",
       y = "Education",
       title = "How does churn risk differ by education")

train %>%
  group_by(income_category) %>%
  summarize_churn() %>%
  plot_categorical(income_category) +
  coord_flip() +
  labs(x = "% churned",
       y = "Income category",
       title = "How does churn risk differ by income category")

qplot(train$customer_age)
train %>%
  group_by(age = cut(customer_age, c(20, 40, 50, 60, 70))) %>%
  summarize_churn() %>%
  plot_categorical(age) +
  coord_flip()
```

Let's look at all numerics!

```{r}
gathered <- train %>%
  select(customer_age, total_relationship_count:churned) %>%
  gather(metric, value, -churned) %>%
  mutate(churned = str_to_title(churned))

gathered %>%
  ggplot(aes(value, fill = churned)) +
  geom_density(alpha = .5) +
  facet_wrap(~ metric, scales = "free") +
  scale_x_log10() +
  labs(title = "How do numeric metrics differ between churn/didn't churn?",
       subtitle = "All predictors on a log scale",
       x = "Log(predictor)",
       fill = "Churned?")

gathered %>%
  group_by(metric) %>%
  mutate(rank = percent_rank(value)) %>%
  ggplot(aes(rank, fill = churned)) +
  geom_density(alpha = .5) +
  facet_wrap(~ metric, scales = "free") +
  labs(title = "How do numeric metrics differ between churn/didn't churn?",
       subtitle = "All predictors on a percentile scale",
       x = "Percentile of predictor",
       fill = "Churned?")
```

Relationships between our important predictors

```{r}
train %>%
  ggplot(aes(total_trans_ct, total_trans_amt, color = churned)) +
  geom_point(alpha = .25) +
  geom_smooth(method = "loess") +
  scale_x_log10() +
  scale_y_log10(labels = dollar_format()) +
  labs(x = "Total transaction count",
       y = "Total transaction amount")

train %>%
  ggplot(aes(total_trans_ct, total_trans_amt / total_trans_ct, color = churned)) +
  geom_point(alpha = .25) +
  geom_smooth(method = "loess") +
  scale_x_log10() +
  scale_y_log10(labels = dollar_format()) +
  labs(x = "Total transaction count",
       y = "Average transaction amount")
```

```{r}
plot_line <- function(tbl, x, ...) {
  tbl %>%
    ggplot(aes({{ x }}, pct_churned, ...)) +
    geom_point(aes(size = n)) +
    geom_line() +
    scale_y_continuous(labels = percent) +
    geom_ribbon(aes(ymin = low, ymax = high), alpha = .5) +
    expand_limits(y = 0) +
    scale_size_continuous(guide = "none") +
    labs(y = "% churned (with 95% confidence bounds)")
}

train %>%
  group_by(total_relationship_count) %>%
  summarize_churn() %>%
  plot_line(total_relationship_count) +
  labs(x = "Total relationships",
       title = "Exploring the importance of # of relationships")
```

The importance of transaction count and transaction amount.

```{r}
# Distribution for cuts
# Two types of products, or two types of customers?
train %>%
  mutate(avg_trans_amt = total_trans_amt / total_trans_ct) %>%
  ggplot(aes(avg_trans_amt)) +
  geom_histogram() +
  geom_vline(xintercept = c(50, 100, 130), lty = 2, color = "red") +
  scale_x_continuous(labels = dollar_format()) +
  labs(title = "Average transaction amount is roughly bimodal",
       subtitle = "Lines shown at 50, 100, 130",
       x = "Average transaction amount")
```

Average transaction amount is roughly bimodal (possibly two types of product, but could also be two types of customer)

```{r}
train %>%
  mutate(avg_trans_amt = total_trans_amt / total_trans_ct) %>%
  group_by(total_trans_ct = cut(total_trans_ct, c(0, 30, 40, 50, 60, 80, Inf)),
           avg_trans_amt = ifelse(avg_trans_amt >= 50, "> $50", "< $50")) %>%
  summarize_churn() %>%
  plot_line(total_trans_ct, group = avg_trans_amt, color = avg_trans_amt) +
  labs(x = "Total transaction count (bucketed)",
       title = "Clients with a moderate number of transactions are most likely to churn",
       color = "Average transaction amount")
```

Customer segments:

* <50 transactions, average > $50: very likely to churn
* <50 transactions, average > $50: moderately likely to churn
* 60+ transactions: very unlikely to churn

(Note that this oversimplifies: the low transaction count are relatively unlikely to churn)

```{r}
train %>%
  mutate(avg_trans_amt = total_trans_amt / total_trans_ct) %>%
  group_by(total_transactions = ifelse(total_trans_ct >= 50,
                                       "> 50 Transactions",
                                       "< 50 Transactions"),
           avg_transaction = ifelse(avg_trans_amt >= 50,
                                    "> $50 Average",
                                    "< $50 Average")) %>%
  summarize_churn() %>%
  ggplot(aes(total_transactions, avg_transaction)) +
  geom_tile(aes(fill = pct_churned)) +
  geom_text(aes(label = percent(pct_churned, 1)), size = 6) +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = .3) +
  labs(x = "How many transactions did the customer do?",
       y = "What was the average transaction size?",
       fill = "% churned",
       title = "Dividing customers into segments based on churn risk")
```

The highest churn rate was among low transaction, high average cost customers!

New features to add:

* The ratio of average transaction amount
* 

