---
title: "NCAA Women's Basketball"
date: 2020-10-06
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2020-10-06")

# Tournament finish - The round of the final game for each team. OR=opening-round loss (1983 only); 1st=first-round loss; 2nd=second-round loss; RSF=loss in the Sweet 16; RF=loss in the Elite Eight; NSF=loss in the national semifinals; N2nd=national runner-up; Champ=national champions
finish_levels <- c("1st", "2nd", "RSF", "RF", "NSF", "N2nd", "Champ")
tournament <- tt$tournament %>%
  filter(year >= 1994) %>%
  mutate(tourney_finish = fct_relevel(tourney_finish, finish_levels),
         n_rounds = as.integer(tourney_finish) - 1)
```

```{r}
tournament %>%
  count(year) %>% 
  ggplot(aes(year, n)) +
  geom_col()
```

```{r}
tournament %>%
  ggplot(aes(full_w, tourney_w + reg_w)) +
  geom_point(alpha = .1)

tournament %>%
  filter(conf_l > reg_l)
```

* Regular wins are before the tournament
* Conf wins are a subset of regular wins
* Tournament wins are in this bracket of 64

There've been 64 teams since 1994.

```{r}
by_seed <- tournament %>%
  group_by(seed) %>%
  summarize(n = n(),
            pct_win = mean(tourney_finish == "Champ"),
            pct_final_four = mean(tourney_finish %in% c("Champ", "N2nd", "NSF")),
            avg_round = mean(as.integer(tourney_finish)))

by_seed %>%
  ggplot(aes(seed, avg_round)) +
  geom_line() +
  labs(x = "Starting seed",
       y = "On average, eliminated in round...")

by_seed %>%
  ggplot(aes(seed, pct_final_four)) +
  geom_line() +
  labs(x = "Starting seed",
       y = "% of times making the Final Four")
```

```{r}
tournament %>%
  count(seed, tourney_finish) %>%
  group_by(seed) %>%
  mutate(pct = n / sum(n)) %>%
  ggplot(aes(tourney_finish, seed, fill = pct)) +
  geom_tile() +
  geom_text(aes(label = percent(pct))) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_reverse(breaks = seq(1, 16), expand = c(0, 0)) +
  scale_fill_gradient2(high = "blue", labels = percent) +
  theme(panel.grid = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "Tournament finish (worst to best)",
       y = "Starting seed",
       fill = "% of seed",
       title = "When a team starts in a seed, how do they end up?")
```

Has the correlation of seededness to # of rounds changed over time?

```{r}
tournament %>%
  ggplot(aes(seed, n_rounds)) +
  geom_point(alpha = .1)

tournament %>%
  group_by(year) %>%
  summarize(correlation = cor(seed, n_rounds)) %>%
  ggplot(aes(year, correlation)) +
  geom_line()
```

No, the predictiveness of a seed hasn't changed in 25 years

Predicting seed + outcome from regular season / conference

```{r}
tournament %>%
  ggplot(aes(seed, reg_percent)) +
  geom_boxplot(aes(group = seed)) +
  geom_smooth(method = "loess")

tournament %>%
  mutate(conference = fct_lump(conference, 8)) %>%
  ggplot(aes(seed, reg_percent)) +
  geom_jitter(width = .05) +
  # geom_boxplot(aes(group = seed)) +
  geom_smooth(method = "loess") +
  facet_wrap(~ conference)
```

```{r}
tournament %>%
  ggplot(aes(reg_percent, n_rounds)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ seed)

tournament %>%
  group_by(seed) %>%
  summarize(correlation = cor(reg_percent, n_rounds))
```

### Conferences

What conferences and schools dominate the NCAA championship?

```{r}
summarize_entries <- . %>%
  summarize(n_entries = n(),
            avg_seed = mean(seed),
            pct_win = mean(tourney_finish == "Champ"),
            pct_final_four = mean(tourney_finish %in% c("Champ", "N2nd", "NSF")),
            avg_round = mean(n_rounds)) %>%
  arrange(desc(n_entries))

by_conference <- tournament %>%
  group_by(conference) %>%
  summarize_entries()

by_conference %>%
  filter(n_entries >= 25) %>%
  ggplot(aes(n_entries, avg_round)) +
  geom_point() +
  geom_text(aes(label = conference), vjust = 1, hjust = 1,
            check_overlap = TRUE) +
  expand_limits(x = 0)
```

```{r}
library(glue)

by_school <- tournament %>%
  group_by(school) %>%
  summarize_entries()

by_school %>%
  filter(n_entries >= 15) %>%
  mutate(school = glue("{ school } ({ n_entries})"),
         school = fct_reorder(school, avg_round)) %>%
  ggplot(aes(avg_round, school)) +
  geom_col() +
  labs(title = "Of schools that frequently play, which perform best?",
       subtitle = "Parentheses show the # of years out of the last 25 they played",
       x = "Average # of rounds won",
       y = "")
```
