---
title: "TidyTemplate"
date: 2020-09-22
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2020-09-22")

peaks <- tt$peaks %>%
  rename(height_meters = height_metres)

peaks %>%
  arrange(desc(height_meters)) %>%
  head(50) %>%
  mutate(peak_name = fct_reorder(peak_name, height_meters)) %>%
  ggplot(aes(height_meters, peak_name, fill = climbing_status)) +
  geom_col() +
  labs(x = "Height (meters)",
       y = "",
       title = "Tallest peaks in the Himalayas",
       fill = "")

na_reasons <- c("Unknown", "Attempt rumoured", "Did not attempt climb", "Did not reach base camp")

expeditions <- tt$expeditions %>%
  mutate(success = case_when(str_detect(termination_reason, "Success") ~ "Success",
                             termination_reason %in% na_reasons ~ "Other",
                             TRUE ~ "Failure")) %>%
  mutate(days_to_highpoint = as.integer(highpoint_date - basecamp_date))
```

* Fraction of successful climbs per mountain, per year
* Rate of death over time / per mountain (by all members or hired members)
* Death rate by mountain and age
* Death causes and rate of injury
* Distribution of length of climbs vs height or vs time
* Correlation between frequency of expeditions and death rate

```{r}
expeditions %>%
  count(termination_reason, sort = TRUE)

expeditions %>%
  filter(!is.na(days_to_highpoint), !is.na(peak_name)) %>%
  filter(success == "Success") %>%
  mutate(peak_name = fct_lump(peak_name, 10),
         peak_name = fct_reorder(peak_name, days_to_highpoint)) %>%
  ggplot(aes(days_to_highpoint, peak_name)) +
  geom_boxplot() +
  labs(x = "Days from basecamp to highpoint",
       y = "",
       title = "How long does it take to get to the high point?",
       subtitle = "Successful climbs only")
```


```{r}
summarize_expeditions <- function(tbl) {
  tbl %>%
    summarize(n_climbs = n(),
              pct_success = mean(success == "Success"),
              across(members:hired_staff_deaths, sum),
              first_climb = min(year)) %>%
    mutate(pct_death = member_deaths / members,
         pct_hired_staff_deaths = hired_staff_deaths / hired_staff)
}

peaks_summarized <- expeditions %>%
  group_by(peak_id, peak_name) %>%
  summarize_expeditions() %>%
  ungroup() %>%
  arrange(desc(n_climbs)) %>%
  inner_join(peaks %>% select(peak_id, height_meters), by = "peak_id")
```

What are the deadliest mountains?

```{r}
# devtools::install_github("dgrtwo/ebbr")
library(ebbr)

peaks_eb <- peaks_summarized %>%
  filter(members >= 20) %>%
  arrange(desc(pct_death)) %>%
  add_ebb_estimate(member_deaths, members)

peaks_eb %>%
  ggplot(aes(pct_death, .fitted)) +
  geom_point(aes(size = members, color = members)) +
  geom_abline(color = "red") +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(labels = percent) +
  scale_color_continuous(trans = "log10") +
  labs(x = "Death rate (raw)",
       y = "Death rate (empirical Bayes adjusted)")

peaks_eb %>%
  filter(members >= 200) %>%
  arrange(desc(.fitted)) %>%
  mutate(peak_name = fct_reorder(peak_name, .fitted)) %>%
  ggplot(aes(.fitted, peak_name)) +
  geom_point(aes(size = members)) +
  geom_errorbarh(aes(xmin = .low, xmax = .high)) +
  expand_limits(x = 0) +
  scale_x_continuous(labels = percent) +
  labs(x = "Death rate (empirical Bayes adjusted + 95% credible interval)",
       y = "",
       title = "How deadly is each peak in the Himalayas?",
       subtitle = "Only peaks that at least 200 climbers have attempted")
```

Recommend: Introduction to Empirical Bayes: Examples from Baseball Statistics

```{r}
# No relationship between height and death rate
peaks_eb %>%
  filter(members >= 100) %>%
  ggplot(aes(height_meters, .fitted)) +
  geom_point(aes(size = members))
```

Explore Everest in particular

```{r}
expeditions %>%
  filter(peak_name == "Everest") %>%
  ggplot(aes(days_to_highpoint, fill = success)) +
  geom_density(alpha = .5)

expeditions %>%
  filter(peak_name == "Everest") %>%
  filter(success == "Success") %>%
  arrange(days_to_highpoint)

expeditions %>%
  filter(success == "Success") %>%
  ggplot(aes(year)) +
  geom_histogram()

everest_by_decade <- expeditions %>%
  filter(peak_name == "Everest") %>%
  mutate(decade = pmax(10 * (year %/% 10), 1970)) %>%
  group_by(decade) %>%
  summarize_expeditions()

everest_by_decade %>%
  ggplot(aes(decade, pct_death)) +
  geom_line(aes(color = "All climbers")) +
  geom_line(aes(y = pct_hired_staff_deaths, color = "Hired staff")) +
  geom_point(aes(color = "All climbers", size = members)) +
  geom_point(aes(y = pct_hired_staff_deaths, size = hired_staff, color = "Hired staff")) +
  scale_x_continuous(breaks = seq(1970, 2010, 10),
                     labels = c("< 1980", seq(1980, 2010, 10))) +
  scale_y_continuous(labels = percent) +
  expand_limits(y = 0) +
  labs(x = "Decade",
       y = "Death rate",
       title = "Everest has been getting less deadly over time",
       subtitle = "Though trends have recently reversed for hired staff",
       size = "# of climbers",
       color = "")
```

### Examine death probability per member

```{r}
members <- tt$members

library(broom)

everest <- members %>%
  filter(peak_name == "Everest")

everest %>%
  group_by(age = 10 * (age %/% 10)) %>%
  summarize(n_climbers = n(),
            pct_death = mean(died))

everest %>%
  group_by(hired) %>%
  summarize(n_climbers = n(),
            pct_death = mean(died))

model <- everest %>%
  mutate(leader = expedition_role == "Leader") %>%
  glm(died ~ year + age + sex + leader + hired + oxygen_used, data = ., family = "binomial")

tidied <- model %>%
  tidy(conf.int = TRUE, exponentiate = TRUE)

tidied %>%
  filter(term != "(Intercept)") %>%
  mutate(term = reorder(term, estimate)) %>%
  ggplot(aes(estimate, term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high))
```

Investigate interaction term of hired + oxygen_used?

