---
title: "TidyTemplate"
date: 2021-01-12
output: html_output
---

# TidyTuesday

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(janitor)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-01-12")

artwork <- tt$artwork %>%
  filter(artist != "Turner, Joseph Mallord William") %>%
  clean_names()

artists <- tt$artists %>%
  clean_names()
```

```{r}
artwork %>%
  separate(medium, c("medium", "surface"), sep = " on ",
           fill = "right", extra = "merge") %>%
  count(medium, sort = TRUE) %>%
  head(20) %>%
  replace_na(list(medium = "Missing")) %>%
  mutate(medium = fct_reorder(medium, n)) %>%
  ggplot(aes(n, medium)) +
  geom_col() +
  labs(y = 'Medium (ignoring "on...")',
       x = "# of pieces in Tate Modern")

artwork %>%
  filter(fct_lump(artist, 16) != "Other") %>%
  mutate(medium = fct_lump(medium, 8)) %>%
  count(artist, medium, sort = TRUE) %>%
  mutate(artist = fct_reorder(artist, n, sum)) %>%
  ggplot(aes(n, artist, fill = medium)) +
  geom_col()
```

```{r}
by_decade_medium <- artwork %>%
  # separate(medium, c("medium", "surface"), sep = " on ",
  #          fill = "right", extra = "merge") %>%
  filter(year >= 1750) %>%
  group_by(decade = round(year, -1)) %>%
  mutate(n_artists_decade = n_distinct(artist)) %>%
  ungroup() %>%
  mutate(medium = fct_lump(medium, 11)) %>%
  group_by(decade, medium, n_artists_decade) %>%
  summarize(n = n(),
            n_artists = n_distinct(artist)) %>%
  ungroup() %>%
  complete(decade, medium, fill = list(n = 0, n_artists = 0)) %>%
  group_by(decade) %>%
  mutate(pct = n / sum(n),
         pct_artists = n_artists / n_artists_decade) %>%
  ungroup()

by_decade_medium %>%
  ggplot(aes(decade, n, fill = medium)) +
  geom_area()

by_decade_medium %>%
  filter(!is.na(medium)) %>%
  mutate(medium = fct_reorder(medium, n, sum)) %>%
  ggplot(aes(decade, pct, fill = medium)) +
  geom_area() +
  scale_y_continuous(labels = percent) +
  labs(x = "Decade",
       y = "% of Tate Modern pieces",
       fill = "Medium (without \"on...\")")

by_decade_medium %>%
  filter(!is.na(medium)) %>%
  mutate(medium = fct_reorder(medium, pct, sum, .desc = TRUE)) %>%
  ggplot(aes(decade, pct, fill = medium)) +
  geom_col() +
  facet_wrap(~ medium, scales = "free_y") +
  scale_y_continuous(labels = percent) +
  theme(legend.position = "none") +
  labs(x = "Decade",
       y = "% of Tate Modern pieces",
       fill = "Medium (without \"on...\")")

by_decade_medium %>%
  filter(!is.na(medium)) %>%
  mutate(medium = fct_reorder(medium, pct_artists, sum, .desc = TRUE)) %>%
  ggplot(aes(decade, pct_artists, fill = medium)) +
  geom_col() +
  facet_wrap(~ medium, scales = "free_y") +
  scale_y_continuous(labels = percent) +
  theme(legend.position = "none") +
  labs(x = "Decade",
       y = "% of artists in this decade using this medium",
       fill = "Medium (without \"on...\")")
```

```{r}
artwork_size <- artwork %>%
  filter(units == "mm",
         !is.na(height),
         !is.na(width)) %>%
  mutate(area = (height / 1000) * (width / 1000),
         ratio = width / height)

artwork_size %>%
  ggplot(aes(area)) +
  geom_histogram() +
  scale_x_log10(labels = comma) +
  labs(x = "Area (square meters)")

artwork_size %>%
  filter(ratio > .2, ratio < 5) %>%
  ggplot(aes(ratio)) +
  geom_histogram(bins = 50) +
  scale_x_log10() +
  geom_vline(xintercept = c(3 / 4, 4 / 3),
             linetype = "dashed", color = "red") +
  labs(x = "Width / height ratio (log scale)",
       title = "What are typical shapes of Tate artwork?",
       subtitle = "Red lines show 3:4 and 4:3 ratios")

artwork_size %>%
  mutate(shape = case_when(
    ratio > 1.05 ~ "Landscape",
    ratio < .95 ~ "Portrait",
    TRUE ~ "Square"
  )) %>%
  count(shape,
        decade = round(year, -1)) %>%
  filter(decade >= 1700) %>%
  complete(decade, shape, fill = list(n = 0)) %>%
  group_by(decade) %>%
  mutate(pct = n / sum(n)) %>%
  ggplot(aes(decade, pct, fill = shape)) +
  geom_area()

size_by_decade <- artwork_size %>%
  group_by(decade = round(year, -1)) %>%
  summarize(median_ratio = median(ratio),
            median_area = median(area),
            n = n()) %>%
  filter(decade >= 1700)

size_by_decade %>%
  ggplot(aes(decade, median_ratio)) +
  geom_line() +
  geom_point(aes(size = n)) +
  labs(x = "Decade",
       y = "Median ratio (width / height)")

size_by_decade %>%
  ggplot(aes(decade, median_area)) +
  geom_line() +
  geom_point(aes(size = n)) +
  labs(x = "Decade",
       y = "Median area in meters^2")

artwork_size %>%
  mutate(decade = round(year, -1)) %>%
  filter(decade >= 1700) %>%
  ggplot(aes(decade, area, group = decade)) +
  geom_boxplot() +
  scale_y_log10(labels = comma,
                breaks = 10 ^ seq(-3, 2))
```

### Artists and size

```{r}
artwork_size %>%
  group_by(artist) %>%
  summarize(n_pieces = n(),
            avg_year = mean(year, na.rm = TRUE),
            first_year = min(year, na.rm = TRUE),
            last_year = max(year, na.rm = TRUE),
            median_area = median(area),
            median_ratio = median(ratio)) %>%
  arrange(desc(n_pieces))

library(glue)

artwork_size %>%
  extract(medium, "medium_on", " on (.*)", remove = FALSE) %>%
  filter(fct_lump(artist, 25) != "Other") %>%
  add_count(artist) %>%
  mutate(artist = glue("{ artist } ({ n })")) %>%
  mutate(artist = fct_reorder(artist, ratio)) %>%
  ggplot(aes(ratio, artist)) +
  geom_boxplot() +
  geom_vline(xintercept = c(3 / 4, 4 / 3),
             linetype = "dashed", color = "red") +
  scale_x_log10(limits = c(.25, 4)) +
  labs(x = "Ratio (width / height)",
       y = "",
       title = "What aspect ratio do artists work in?",
       subtitle = "For the 25 most common artists in the Tate")

library(ggridges)
artwork_size %>%
  extract(medium, "medium_on", " on (.*)", remove = FALSE) %>%
  filter(fct_lump(artist, 25) != "Other") %>%
  add_count(artist) %>%
  mutate(artist = glue("{ artist } ({ n })")) %>%
  mutate(artist = fct_reorder(artist, ratio)) %>%
  ggplot(aes(ratio, artist)) +
  geom_density_ridges() +
  geom_vline(xintercept = c(3 / 4, 4 / 3),
             linetype = "dashed", color = "red") +
  scale_x_log10(limits = c(.33, 3)) +
  labs(x = "Ratio (width / height)",
       y = "",
       title = "What aspect ratio do artists work in?",
       subtitle = "For the 25 most common artists in the Tate")

artwork_size %>%
  filter(fct_lump(medium, 25) != "Other") %>%
  add_count(medium) %>%
  mutate(medium = glue("{ medium } ({ n })")) %>%
  mutate(medium = fct_reorder(medium, ratio)) %>%
  ggplot(aes(ratio, medium)) +
  geom_boxplot() +
  geom_vline(xintercept = c(3 / 4, 4 / 3),
             linetype = "dashed", color = "red") +
  scale_x_log10(limits = c(.25, 4)) +
  labs(x = "Ratio (width / height)",
       y = "",
       title = "What aspect ratio is a medium typically in?",
       subtitle = "For the 25 most common medium in the Tate")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}



```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}


  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
