---
title: "Student/Teacher Ratios"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
theme_set(theme_light())

student_ratio <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-07/student_teacher_ratio.csv")
```

```{r}
student_teacher_ratio_2015 <- student_ratio %>%
  filter(indicator == "Primary Education",
         year == 2015,
         !is.na(student_ratio))

student_teacher_ratio_2015 %>%
  arrange(desc(student_ratio)) %>%
  slice(c(1:10, seq(n() - 10, n()))) %>%
  mutate(country = fct_reorder(country, student_ratio)) %>%
  ggplot(aes(country, student_ratio)) +
  geom_point() +
  coord_flip() +
  expand_limits(y = 0) +
  labs(title = "Countries with the highest and lowest student/teacher ratios",
       x = "",
       y = "Student/teacher ratio")
```

It looks based on my knowledge of geography that student/teacher ratio is negatively correlated with a country's wealth: some of the richest countries TODO. But there are clear exceptions: some rich countries like the US, UK, and Switzerland don't appear in the low student/teacher ratios, while relatively lower-income countries like Cuba, Greece and Georgia do. We'll keep an eye out for those and other outliers.

```{r}
library(WDI)

WDIsearch("public.*education") %>%
  as.data.frame() %>%
  tbl_df() %>%
  arrange(str_length(name)) %>%
  View()
```

```{r}
indicators_raw <- WDI(indicator = c("NY.GDP.PCAP.CD", "SP.POP.TOTL", "SE.ADT.LITR.ZS",
                                    "SE.XPD.TOTL.GD.ZS",
                                    "SE.SEC.NENR.MA", "SE.SEC.NENR.FE"),
                      start = 2015, end = 2015, extra = TRUE) %>%
  tbl_df()

indicators <- indicators_raw %>%
  select(country_code = iso3c,
         region,
         NY.GDP.PCAP.CD:SE.SEC.NENR.FE) %>%
  mutate(country_code = as.character(country_code))

student_teacher_ratio_2015 %>%
  ggplot(aes(student_ratio)) +
  geom_histogram() +
  scale_x_log10()
```

```{r}
student_teacher_ratio_2015 %>%
  inner_join(indicators, by = "country_code") %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.CD, student_ratio)) +
  geom_point(aes(size = SP.POP.TOTL, color = region)) +
  geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_x_log10() +
  scale_y_log10() +
  scale_size_continuous(labels = scales::comma_format(), range = c(.25, 12)) +
  labs(x = "GDP per capita",
       y = "Student/teacher ratio in primary education",
       subtitle = "In 2015",
       title = "GDP per capita and student/teacher ratio are negatively correlated",
       color = "Region",
       size = "Population")
```

This confirms there's a negative correlation between a country's wealth and its student/teacher ratio.

### Other indicators

```{r}
student_teacher_ratio_2015 %>%
  inner_join(indicators, by = "country_code") %>%
  filter(!is.na(SE.ADT.LITR.ZS)) %>%
  mutate(literacy = SE.ADT.LITR.ZS / 100) %>%
  ggplot(aes(student_ratio, literacy)) +
  geom_point() +
  geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_x_log10() +
  scale_y_continuous(labels = scales::percent_format())
```


```{r}
joined <- student_teacher_ratio_2015 %>%
  inner_join(indicators, by = "country_code") %>%
  mutate(secondary_enrollment = (SE.SEC.NENR.MA + SE.SEC.NENR.FE) / 2)

joined %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.CD, secondary_enrollment)) +
  geom_point() +
  geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_x_log10() +
  labs(x = "GDP per capita",
       y = "Secondary school enrollment")
```

```{r}
joined %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(student_ratio, secondary_enrollment)) +
  geom_point() +
  geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE) +
  labs(x = "Student/teacher ratio in primary school",
       y = "Secondary school enrollment")
```

```{r}
joined %>%
  transmute(student_ratio,
            secondary_enrollment,
            log2_gdp = log2(NY.GDP.PCAP.CD)) %>%
  cor(use = "pairwise.complete.obs")
```


Examining the confounding variable:

```{r}
joined %>%
  lm(secondary_enrollment ~ student_ratio + log2(NY.GDP.PCAP.CD),
     data = .) %>%
  summary()
```




### Appendix: Primary vs secondary education

```{r}
secondary_primary_education <- student_ratio %>%
  filter(year == 2015,
         !is.na(student_ratio),
         indicator %in% c("Primary Education", "Secondary Education")) %>%
  group_by(country) %>%
  filter(n() == 2) %>%
  ungroup()

secondary_primary_education %>%
  inner_join(indicators, by = "country_code") %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.CD, student_ratio)) +
  geom_point(aes(size = SP.POP.TOTL, color = region)) +
  geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_x_log10() +
  scale_y_log10() +
  scale_size_continuous(labels = scales::comma_format(), range = c(.25, 12)) +
  facet_wrap(~ indicator) +
  labs(x = "GDP per capita",
       y = "Student/teacher ratio in primary education",
       subtitle = "In 2015",
       title = "GDP per capita and student/teacher ratio are negatively correlated",
       color = "Region",
       size = "Population")

secondary_primary_education %>%
  select(indicator, country, student_ratio) %>%
  mutate(indicator = snakecase::to_snake_case(indicator)) %>%
  spread(indicator, student_ratio) %>%
  mutate(ratio = secondary_education / primary_education) %>%
  ggplot(aes(primary_education, secondary_education)) +
  geom_point() +
  geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_x_log10() +
  scale_y_log10()
```


```{r}

```






