---
title: "TidyTemplate"
date: 2020-08-25
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-08-25")

tt

chopped <- tt$chopped

```

```{r}
chopped %>%
  ggplot(aes(episode_rating)) +
  geom_histogram()

chopped %>%
  arrange(episode_rating) %>%
  View()

chopped %>%
  filter(!is.na(episode_rating)) %>%
  ggplot(aes(series_episode, episode_rating)) +
  geom_line(alpha = .5, color = "gray") +
  geom_point(aes(color = factor(season))) +
  geom_text(aes(label = episode_name), hjust = 1,
            check_overlap = TRUE) +
  theme(legend.position = "none")

chopped %>%
  filter(!is.na(episode_rating)) %>%
  group_by(season) %>%
  summarize(n_episodes = n(),
            avg_rating = mean(episode_rating)) %>%
  ggplot(aes(season, avg_rating)) +
  geom_line() +
  geom_point(aes(size = n_episodes)) +
  theme(legend.position = "none") +
  labs(x = "Season",
       y = "Average Rating")
```

```{r}
library(glue)

chopped %>%
  arrange(desc(episode_rating)) %>%
  head(25) %>%
  mutate(name = glue("{ season }.{season_episode} { episode_name }"),
         name = fct_reorder(name, episode_rating)) %>%
  ggplot(aes(episode_rating, name)) +
  geom_point()
```

### Ingredients!

```{r}
ingredients <- chopped %>%
  select(season, season_episode, series_episode, episode_name,
         episode_rating, appetizer:dessert) %>%
  pivot_longer(cols = c(appetizer:dessert),
               names_to = "course",
               values_to = "ingredient") %>%
  separate_rows(ingredient, sep = ", ") %>%
  mutate(course = fct_relevel(course, c("appetizer", "entree")))

ingredients %>%
  count(course, ingredient, sort = TRUE) %>%
  filter(fct_lump(ingredient, 25, w = n) != "Other") %>%
  mutate(ingredient = fct_reorder(ingredient, n, sum),
         course = fct_rev(course)) %>%
  ggplot(aes(n, ingredient, fill = course)) +
  geom_col() +
  scale_fill_discrete(guide = guide_legend(reverse = TRUE)) +
  labs(x = "# of episodes",
       y = "",
       title = "Most common ingredients in Chopped",
       fill = "Course")
```

TODO: Link to slides from widyr talk

```{r}
library(widyr)
library(ggraph)
library(tidygraph)

ingredients_filtered <- ingredients %>%
  add_count(ingredient) %>%
  filter(n >= 8)

ingredient_correlations <- ingredients_filtered %>%
  pairwise_cor(ingredient, series_episode, sort = TRUE)

ingredients_filtered %>%
  pairwise_count(ingredient, series_episode, sort = TRUE)

# Not sure this is useful since they appear across courses
ingredient_correlations %>%
  head(75) %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation)) +
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE)

# Do any pairs of ingredients appear together in the same course
# across episodes?
ingredients_filtered %>%
  unite(episode_course, series_episode, course) %>%
  pairwise_count(ingredient, episode_course, sort = TRUE)
```

# Are there ingredients that got more popular in later seasons?

```{r}
early_late_ingredients <- ingredients_filtered %>%
  group_by(ingredient) %>%
  summarize(first_season = min(season),
            avg_season = mean(season),
            last_season = max(season),
            n_appearances = n()) %>%
  arrange(desc(avg_season)) %>%
  slice(c(1:6, tail(row_number())))

ingredients_filtered %>%
  semi_join(early_late_ingredients, by = "ingredient") %>%
  mutate(ingredient = fct_reorder(ingredient, season)) %>%
  ggplot(aes(season, ingredient)) +
  geom_boxplot()
```

### What ingredients are "popular"

Which ingredients lead to an above- or below- average episode?

```{r}
ingredients_wide <- ingredients_filtered %>%
  select(season, series_episode, episode_rating, ingredient) %>%
  mutate(value = 1) %>%
  pivot_wider(names_from = ingredient,
              values_from = value,
              values_fill = list(value = 0)) %>%
  select(-series_episode) %>%
  janitor::clean_names()

lm(episode_rating ~ season, data = ingredients_wide) %>%
  summary()

library(tidymodels)

set.seed(2020)
split_data <- ingredients_wide %>%
  filter(!is.na(episode_rating)) %>%
  initial_split()

training_set <- training(split_data)
```

```{r}
cv_samples <- training_set %>%
  vfold_cv(v = 10)
```

```{r}
model_spec <- linear_reg(penalty = tune()) %>%
  set_engine("glmnet")

parameter_search <- model_spec %>%
  tune_grid(episode_rating ~ ., resamples = cv_samples)

parameter_search %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  ggplot(aes(penalty, mean)) +
  geom_line() +
  scale_x_log10() +
  labs(y = "Mean Squared Error")
```

```{r}
rf_spec <- rand_forest(mode = "regression",
                       trees = tune()) %>%
  set_engine("ranger")

cv_samples <- training_set %>%
  vfold_cv(v = 10)

parameter_search <- rf_spec %>%
  tune_grid(episode_rating ~ ., resamples = cv_samples)

parameter_search %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  ggplot(aes(trees, mean)) +
  geom_line() +
#   scale_x_log10() +
  labs(x = "# of trees in random forest",
       y = "Mean Squared Error")

model <- rand_forest(mode = "regression", mtry = 3, trees = 500) %>%
  set_engine("ranger") %>%
  fit(episode_rating ~ ., training_set)

test_set <- testing(split_data)

predict(model, test_set) %>%
  bind_cols(test_set) %>%
  rmse(.pred, episode_rating)

test_set %>%
  mutate(average = mean(episode_rating)) %>%
  rmse(average, episode_rating)
```

Note to self: lower RMSE is better, go back to college Dave.

TODO: bug report for one predictor in glmnet engine linear_reg

Conclusion: can't improve on a dummy model for predicting episode quality from ingredients

Spline model

```{r}
rec <- recipe(episode_rating ~ season, training_set) %>%
  step_ns(season, deg_free = tune())

parameter_search_df <- linear_reg() %>%
  set_engine("lm") %>%
  tune_grid(rec, resamples = cv_samples)

parameter_search_df %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  ggplot(aes(deg_free, mean)) +
  geom_line()

training_data_processed <- recipe(episode_rating ~ season, training_set) %>%
  step_ns(season, deg_free = 2) %>%
  prep() %>%
  juice()

spline_model <- linear_reg() %>%
  set_engine("lm") %>%
  fit(episode_rating ~ season, data = juice(training_data_processed))
```

