---
title: "TidyTemplate"
date: 2021-03-09
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-03-09")

movies <- tt$movies %>%
  mutate(across(contains("gross"), parse_number))

View(movies)
```

```{r Readme, eval = interactive()}
movies %>%
  count(year, binary) %>%
  ggplot(aes(year, n, fill = binary)) +
  geom_col()

summarize_bechdel <- function(tbl) {
  tbl %>%
    summarize(n_movies = n(),
              pct_pass = mean(binary == "PASS"),
              median_intgross_2013 = median(intgross_2013, na.rm = TRUE)) %>%
    arrange(desc(n_movies))
}

movies %>%
  group_by(decade = 10 * (year %/% 10)) %>%
  summarize_bechdel() %>%
  ggplot(aes(decade, pct_pass)) +
  geom_line() +
  geom_point(aes(size = n_movies)) +
  expand_limits(y = 0)

movies %>%
  arrange(desc(intgross_2013)) %>%
  head(25) %>%
  mutate(title = fct_reorder(title, intgross_2013)) %>%
  ggplot(aes(intgross_2013, title, fill = binary)) +
  geom_col()

movies %>%
  ggplot(aes(intgross_2013, binary)) +
  geom_boxplot() +
  scale_x_log10()

movies %>%
  filter(year >= 1990) %>%
  ggplot(aes(intgross_2013, fill = binary)) +
  geom_density(alpha = .5) +
  scale_x_log10()

lm(log2(intgross_2013) ~ binary, data = movies) %>%
  summary()
```

```{r}
library(ggrepel)

movie_genres <- movies %>%
  separate_rows(genre, sep = ", ") %>%
  filter(year >= 1990, !is.na(genre))

movie_genres %>%
  group_by(genre) %>%
  summarize_bechdel() %>%
  head(12) %>%
  ggplot(aes(median_intgross_2013, pct_pass)) +
  geom_point(aes(size = n_movies)) +
  geom_text_repel(aes(label = genre)) +
  expand_limits(x = 0, y = 0) +
  scale_y_continuous(labels = percent) +
  labs(x = "Median International Gross (2013 dollars)",
       y = "% that pass the Bechdel test")

movie_genres %>%
  filter(fct_lump(genre, 9) != "Other") %>%
  ggplot(aes(intgross_2013, fill = binary)) +
  geom_density(alpha = .5) +
  facet_wrap(~ genre) +
  scale_x_log10(labels = dollar) +
  labs(x = "International Gross (2013 dollars)")

movies %>%
  filter(fct_lump(genre, 9) != "Other") %>%
  ggplot(aes(intgross_2013, genre, fill = binary)) +
  geom_boxplot() +
  scale_x_log10(labels = dollar) +
  labs(x = "International Gross (2013 dollars)")

movies

movies %>%
  filter(year >= 1990) %>%
  separate_rows(genre, sep = ", ") %>%
  mutate(value = 1) %>%
  spread(genre, value, fill = 0) %>%
  lm(log2(intgross_2013) ~ Action + Adventure + Crime +
       Thriller + `Sci-Fi` + Fantasy + Comedy +
       Drama + Romance + Horror +
       binary, data = .) %>%
  summary()
```

```{r}
genre_reasons <- movie_genres %>%
  count(genre = fct_lump(genre, 8),
        clean_test,
        sort = TRUE) %>%
  group_by(genre) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

genre_reasons %>%
  ggplot(aes(pct, clean_test)) +
  geom_col() +
  facet_wrap(~ genre) +
  scale_x_continuous(labels = percent) +
  labs(title = "Does the reason a movie fails the Bechdel test differ between genres?",
       y = "Test result",
       x = "% of movies in this genre")

library(tidytext)

genre_reasons %>%
  filter(clean_test != "dubious") %>%
  mutate(clean_test = fct_relevel(clean_test, "nowomen", "notalk", "men"),
         clean_test = fct_recode(clean_test,
                                 "Passes!" = "ok",
                                 "Has two women..." = "nowomen",
                                 "...who talk to each other..." = "notalk",
                                 "...about something other than a man" = "men")) %>%
  mutate(genre = reorder_within(genre, pct, clean_test)) %>%
  ggplot(aes(pct, genre)) +
  geom_col() +
  facet_wrap(~ clean_test, scales = "free_y") +
  scale_x_continuous(labels = percent) +
  scale_y_reordered() +
  labs(title = "Which step do movies fail on?",
       y = "",
       x = "% of movies in this genre that fail at this step?")
```

```{r}
movies %>%
  filter(!is.na(director)) %>%
  group_by(director = fct_lump(director, 12)) %>%
  summarize_bechdel() %>%
  arrange(desc(pct_pass))

movies %>%
  filter(director == "Steven Spielberg") %>%
  arrange(desc(year)) %>%
  View()

movies %>%
  filter(!is.na(writer)) %>%
  separate_rows(writer, sep = ", ") %>%
  mutate(writer = str_remove(writer, " \\(.*")) %>%
  distinct(title, writer, .keep_all = TRUE) %>%
  group_by(writer = fct_lump(writer, 10)) %>%
  summarize_bechdel() %>%
  View()
```


```{r}
movie_words <- movies %>%
  filter(!is.na(plot)) %>%
  unnest_tokens(word, plot) %>%
  anti_join(stop_words, by = "word")

movie_words %>%
  distinct(word, title, .keep_all = TRUE) %>%
  group_by(word = fct_lump(word, 50)) %>%
  summarize_bechdel() %>%
  arrange(desc(pct_pass)) %>%
  view()
```

glmnet

```{r}
library(glmnet)

movie_features <- bind_rows(
  movie_words %>% distinct(imdb, title, binary, feature = word),
  movie_genres %>% distinct(imdb, title, binary, feature = genre)
) %>%
  add_count(feature) %>%
  filter(n >= 10)

movie_feature_matrix <- movie_features %>%
  cast_sparse(imdb, feature)

pass <- movies$binary[match(rownames(movie_feature_matrix), movies$imdb)] == "PASS"

mod <- cv.glmnet(movie_feature_matrix, pass, family = "binomial")

plot(mod)

library(broom)

tidy(mod$glmnet.fit) %>%
  filter(lambda == mod$lambda.1se,
         term != "(Intercept)") %>%
  top_n(40, abs(estimate)) %>%
  mutate(term = fct_reorder(term, estimate)) %>%
  ggplot(aes(estimate, term)) +
  geom_col() +
  labs(y = "Word/Genre",
       x = "Coefficient towards passing Bechdel test")
```

# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}



```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}


  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
