---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(scales)
library(lubridate)
library(textrecipes)
library(stacks)
theme_set(theme_light())

set.seed(2021)
```

```{r}
dataset <- read_csv("~/Downloads/sliced-s01e04-knyna9/train.csv") %>%
  mutate(rain_tomorrow = factor(ifelse(rain_tomorrow, "Rained", "Didn't Rain")))
holdout <- read_csv("~/Downloads/sliced-s01e04-knyna9/test.csv") %>%
  mutate(rain_tomorrow = factor(ifelse(rain_tomorrow, "Rained", "Didn't Rain")))

set.seed(2021)
spl <- initial_split(dataset, prop = .75)
train <- training(spl)
test <- testing(spl)

train_fold <- train %>%
  vfold_cv(v = 5)

doParallel::registerDoParallel(cores = 4)doPrimitiveMethod()
```

```{r}
mset <- metric_set(mn_log_loss)

grid_control <- control_grid(save_pred = TRUE,
                             save_workflow = TRUE,
                             extract = extract_model)

augment.workflow <- function(x, newdata, ...) {
  predict(x, newdata, ...) %>%
    bind_cols(newdata)
}
augment.model_stack <- function(x, newdata, ...) {
  predict(x, newdata, ...) %>%
    bind_cols(newdata)
}


predict_on_holdout <- function(workflow) {
  workflow %>%
    fit(dataset) %>%
    augment(holdout, type = "prob") %>%
    select(id, rain_tomorrow = .pred_Rained)
}
```

```{r}
summarize_rainfall <- function(tbl) {
  ret <- tbl %>%
    summarize(n_rain = sum(rain_tomorrow == "Rained"),
              n = n()) %>%
    arrange(desc(n)) %>%
    ungroup() %>%
    mutate(pct_rain = n_rain / n,
           low = qbeta(.025, n_rain + 5, n - n_rain + .5),
           high = qbeta(.975, n_rain + 5, n - n_rain + .5)) %>%
    mutate(pct = n / sum(n))

  ret
}

train %>%
  summarize_rainfall()

train %>%
  group_by(location = fct_lump(location, 50)) %>%
  summarize_rainfall() %>%
  mutate(location = fct_reorder(location, pct_rain)) %>%
  ggplot(aes(pct_rain, location)) +
  geom_point(aes(size = pct)) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = .3) +
  scale_size_continuous(labels = percent, guide = "none", range = c(.5, 4)) +
  scale_x_continuous(labels = percent) +
  labs(x = "Probability of raining tomorrow",
       y = "",
       title = "What locations get the most/least rain?",
       subtitle = "Including 95% confidence intervals. Size of points is proportional to frequency")
```

Time series! By year and within year

```{r}
train %>%
  group_by(year = year(date)) %>%
  summarize_rainfall() %>%
  ggplot(aes(year, pct_rain)) +
  geom_point(aes(size = n)) +
  geom_line() +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = .1) +
  expand_limits(y = 0) +
  scale_x_continuous(breaks = seq(2007, 2017, 2)) +
  scale_y_continuous(labels = percent_format()) +
  scale_size_continuous(guide = "none") +
  coord_cartesian(ylim = c(0, .35)) +
  labs(x = "Year",
       y = "% days rain (tomorrow) in year")
```

```{r}
numeric_gathered <- train %>%
  mutate(rainfall = log2(rainfall + 1)) %>%
  gather(metric, value, min_temp, max_temp, rainfall, contains("speed"), contains("humidity"), contains("pressure"), contains("cloud"), contains("temp"))

numeric_gathered %>%
  ggplot(aes(value, color = rain_tomorrow)) +
  geom_density() +
  facet_wrap(~ metric, scales = "free")
```

Lots of linear trends! How well would each of these do by itself (AUC)?

```{r}
numeric_gathered %>%
  group_by(metric) %>%
  roc_auc(rain_tomorrow, value, event_level = "second") %>%
  arrange(desc(.estimate)) %>%
  mutate(metric = fct_reorder(metric, .estimate)) %>%
  ggplot(aes(.estimate, metric)) +
  geom_point() +
  geom_vline(xintercept = .5) +
  labs(x = "AUC in positive direction",
       title = "How predictive is each linear predictor by itself?",
       subtitle = ".5 is not predictive at all; <.5 means negatively associated with rain, >.5 means positively associated")

```

```{r}
train %>%
  group_by(wind_gust_dir = str_sub(wind_gust_dir, 1, 1)) %>%
  summarize_rainfall()
```


It looks like the ones with the strongest trends are

There's a bit of a time trend maybe. It's nonlinear and not huge.

```{r}
train %>%
  group_by(week = as.Date("2020-01-01") + week(date) * 7) %>%
  summarize_rainfall() %>%
  ggplot(aes(week, pct_rain)) +
  geom_point(aes(size = n)) +
  geom_line() +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = .1) +
  expand_limits(y = 0) +
  scale_x_date(labels = date_format("%b")) +
  scale_y_continuous(labels = percent) +
  scale_size_continuous(guide = "none") +
  labs(x = "Week within year",
       y = "% days rain (tomorrow)",
       title = "Rain really is more common in the summer!",
       subtitle = "Ribbon shows 95% confidence bound")
```

Compass rose

```{r}
compass_directions <- c("N", "NNE", "NE", "ENE", "E", "ESE", "SE", "SSE", "S",
  "SSW", "SW", "WSW", "W", "WNW", "NW", "NNW")

as_angle <- function(direction) {
  (match(direction, compass_directions) - 1) * 360 / length(compass_directions)
}
```

```{r}
train %>%
  gather(type, direction, contains("dir")) %>%
  mutate(type = str_remove(type, "wind_dir"),
         type = ifelse(type == "wind_gust_dir", "Overall", type)) %>%
  group_by(type, direction) %>%
  summarize_rainfall() %>%
  mutate(angle = as_angle(direction)) %>%
  ggplot(aes(angle)) +
  geom_line(aes(y = pct_rain, color = type)) +
  geom_text(aes(y = .4, label = direction), size = 6) +
  scale_y_continuous(labels = percent) +
  expand_limits(y = 0) +
  coord_polar() +
  labs(x = "",
       y = "% rain tomorrow when wind is in this direction",
       title = "How does rain depend on the compass direction?",
       color = "Time")
```


Looking for interaction terms in change over day!

```{r}
spread_time <- train %>%
  select(id, rain_tomorrow, contains("3"), contains("9")) %>%
  select(-contains("dir")) %>%
  gather(metric, value, -rain_tomorrow, -id) %>%
  separate(metric, c("metric", "time"), sep = -3) %>%
  mutate(time = paste0("at_", time)) %>%
  spread(time, value)

spread_time %>%
  ggplot(aes(at_9am, at_3pm)) +
  geom_point() +
  geom_smooth(aes(color = rain_tomorrow), method = "lm") +
  facet_wrap(~ metric, scales = "free") +
  labs(x = "9 AM",
       y = "3 PM",
       title = "Does the difference over the course of the day differ by rain/won't rain?",
       color = "")
```

```{r}
spread_time %>%
  ggplot(aes(at_3pm - at_9am)) +
  geom_density(aes(color = rain_tomorrow)) +
  facet_wrap(~ metric, scales = "free") +
  labs(x = "Difference between 3 PM - 9 AM",
       title = "Does the difference over the course of the day differ by rain/won't rain?",
       color = "")
```

Conclusions:

* If humidity increased, likely to rain
* If pressure increased, slightly likely to rain
* If temp increased, less likely to rain
* No effect of change in wind speed or cloud cover from 9am to 3pm


```{r}
prep_juice <- function(x) juice(prep(x))

lin_rec <- recipe(rain_tomorrow ~ date + location + rain_today +
                    min_temp + max_temp + rainfall +
                    wind_gust_speed + wind_speed9am +
                    wind_speed3pm + humidity9am + humidity3pm + pressure9am +
                    pressure3pm + cloud9am + cloud3pm + temp9am + temp3pm +
                    wind_gust_dir + wind_dir9am + wind_dir3pm +
                    rain_today, data = train) %>%
  step_mutate(year = year(date),
              month = month(date)) %>%
  step_rm(date) %>% 
  step_log(rainfall, offset = 1, base = 2) %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_ns(year, deg_free = 2) %>%
  step_ns(month, deg_free = 4) %>%
  step_other(location, threshold = .005) %>%
  step_dummy(location) %>%
  step_mutate(wind_gust_dir = str_sub(wind_gust_dir, 1, 1)) %>%
  step_mutate(wind_dir9am = str_sub(wind_dir9am, 1, 1)) %>%
  step_mutate(wind_dir3pm = str_sub(wind_dir3pm, 1, 1)) %>%
  step_unknown(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())

lin_wf <- workflow() %>%
  add_recipe(lin_rec) %>%
  add_model(logistic_reg(penalty = tune()) %>% set_engine("glmnet"))

lin_tune <- lin_wf %>%
  tune_grid(train_fold,
            grid = crossing(penalty = 10 ^ seq(-6, -.5, .05)),
            metrics = mset,
            control = grid_control)

autoplot(lin_tune)
```

First model!

```{r}
lin_wf_best <- lin_wf %>%
  finalize_workflow(select_best(lin_tune))

lin_wf_best %>%
  last_fit(spl, metrics = mset) %>%
  collect_metrics()

lin_wf_best %>%
  predict_on_holdout() %>%
  write_csv("~/Desktop/attempt1.csv")
```

XGBoost

Trying out one more xgboost model: adding a location term, and tuning how many we include.

```{r}
set.seed(2021)

train_fold_small <- train %>%
  sample_n(4000) %>%
  vfold_cv(v = 2)

compass_directions <- c("N", "NNE", "NE", "ENE", "E", "ESE", "SE", "SSE", "S",
  "SSW", "SW", "WSW", "W", "WNW", "NW", "NNW")

as_angle <- function(direction) {
  (match(direction, compass_directions) - 1) * 360 / length(compass_directions)
}


xg_rec <- recipe(rain_tomorrow ~ date + rain_today +
                    min_temp + max_temp + rainfall +
                    wind_gust_speed + wind_speed9am +
                    wind_speed3pm + humidity9am + humidity3pm + pressure9am +
                    pressure3pm + cloud9am + cloud3pm + temp9am + temp3pm +
                    wind_gust_dir + wind_dir9am + wind_dir3pm +
                    location +
                    rain_today, data = train) %>%
  step_mutate(week = week(date)) %>%
  step_mutate(wind_gust_dir = as_angle(wind_gust_dir)) %>%
  step_mutate(wind_dir9am = as_angle(wind_dir9am)) %>%
  step_mutate(wind_dir3pm = as_angle(wind_dir3pm)) %>%
  step_other(location, threshold = tune()) %>%
  step_rm(date) %>% 
  step_impute_mean(all_numeric_predictors()) %>%
  step_unknown(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())

xg_wf <- workflow() %>%
  add_recipe(xg_rec) %>%
  add_model(boost_tree("classification",
                       mtry = tune(),
                       trees = tune(),
                       learn_rate = .01) %>% set_engine("xgboost"))

xg_tune <- xg_wf %>%
  tune_grid(train_fold,
            grid = crossing(mtry = c(9, 12),
                       trees = seq(100, 1600, 100),
                       threshold = c(.01)),
            metrics = mset,
            control = grid_control)

autoplot(xg_tune)
```

```{r}
xg_wf_best <- xg_wf %>%
  finalize_workflow(select_best(xg_tune))

xg_fit_best <- xg_wf_best %>%
  fit(train)

importances <- xgboost::xgb.importance(model = xg_fit_best$fit$fit$fit)

importances %>%
  mutate(Feature = fct_reorder(Feature, Gain)) %>%
  ggplot(aes(Gain, Feature)) +
  geom_point() +
  labs(title = "Importance of each term in xgboost",
       subtitle = "Even after turning direction numeric, still not *that* important")

predictions <- xg_fit_best %>%
  augment(test, type = "prob")

predictions %>%
  mn_log_loss(rain_tomorrow, `.pred_Didn't Rain`)
```

```{r}
xg_wf_best %>%
  predict_on_holdout() %>%
  write_csv("~/Desktop/attempt3.csv")
```

So far, xgboost slightly better .349 with gust turned numeric

TODO:

* Blend linear and xgboost (to get benefit of location)
* Relationships between predictors


```{r}
prep_juice <- function(x) juice(prep(x))

lin_location_rec <- recipe(rain_tomorrow ~ date + rainfall + location + humidity3pm, data = train) %>%
  step_mutate(week = week(date)) %>%
  step_rm(date) %>% 
  step_log(rainfall, offset = 1, base = 2) %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_ns(week, deg_free = 4) %>%
  step_other(location, threshold = .005) %>%
  step_unknown(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())

lin_location_wf <- workflow() %>%
  add_recipe(lin_location_rec) %>%
  add_model(logistic_reg(penalty = tune()) %>% set_engine("glmnet"))

lin_location_tune <- lin_location_wf %>%
  tune_grid(train_fold,
            grid = crossing(penalty = 10 ^ seq(-6, -.5, .05)),
            metrics = mset,
            control = grid_control)

autoplot(lin_location_tune)
```

```{r}
lin_best <- lin_location_tune %>% filter_parameters(parameters = select_best(lin_location_tune))
xg_best <- xg_tune %>% filter_parameters(parameters = select_best(xg_tune))

lin_xg_blend <- stacks() %>%
  add_candidates(lin_best) %>%
  add_candidates(xg_best) %>%
  blend_predictions()

lin_xg_fit <- lin_xg_blend %>%
  fit_members()

# Performance on test
predictions <- lin_xg_fit %>%
  predict(test, type = "prob", members = TRUE)

# Log loss by model, or by the blend
predictions %>%
  select(contains("_Rained")) %>%
  bind_cols(select(test, rain_tomorrow)) %>%
  gather(model, prediction, -rain_tomorrow) %>%
  mutate(prediction = 1 - prediction) %>%
  group_by(model) %>%
  mn_log_loss(rain_tomorrow, prediction)

  mn_log_loss(rain_tomorrow, `.pred_Didn't Rain`)
```


### Old stuff

Let's look at other relations

Predictors:

Year (weak), week within year (mostly summer)
Location: strong effect and fairly evenly spread out
Most important: pressure, humidity, cloud, temp3pm. 


Compass rose plot on polar coordinates (while I'm waiting for xgboost, haha)


